---
title: "Cleaning and Wrangling"
author: "Tobias M. Osborne"
format: html
---

## Set up

```{r}
#| message: false
#| warning: false

library(dplyr)
library(tidyr)
library(readr)

```

## Read Data

```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

## Explore Data

```{r}

colnames(catch_original)

head(catch_original)

summary(catch_original)

unique(catch_original$Region)

```

## Clean and Reshape

### Use `select()`

```{r}
catch_data <- catch_original %>%
  select(Region,Year,Chinook, Sockeye,Coho, Pink, Chum)

colnames(catch_data)
```

### Quality Check

```{r}

glimpse(catch_data)
```


### Replace values in column using `mutate()`

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = as.numeric(Chinook))
```

finding rows with NA values

```{r}
i <- which(is.na(catch_clean$Chinook))

catch_data[i,]
```

Fixing the i in Chinook column
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(condition = (Chinook == "I"), 
                           true = "1",
                           false = Chinook),
         Chinook = as.numeric(Chinook))
```

### Re-shape data with `pivot_longer()` and `pivot_winder`

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region,Year),
               names_to = "species",
               values_to = "catch")
```

Reversing from long to wide
```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = species,
              values_from = catch)

head(catch_wide)
```

### Rename columne using `rename()`

```{r}
catch_long <- catch_long %>% 
  rename(catch_thousand = catch)

head(catch_long)
```

### Add a New column using `mutate()`

```{r}
catch_long <- catch_long %>% 
  mutate(catch = catch_thousand*1000) %>% 
  select(-catch_thousand)

head(catch_long)
```

### Demostrate Split-Apply_Combine

```{r}
mean_region <- catch_long %>%
  group_by(Region) %>%
  summarize(mean_catch = mean(catch))

head(mean_region)

n_region <- catch_long %>%
  group_by(Region) %>%
  summarise(n = n())

head(n_region)
```


### Use `filter()` to subset by rows

```{r}

sse_cacth <- catch_long %>% 
  filter(Region == "SSE")

unique(sse_cacth$Region)
head(sse_cacth)
```


### Sort Data from using `arrange()`

### Split and Combin 

```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAU-039"))

sites_df %>%
    separate(site, c("island", "site_number"), "-")


```

```{r}
dates_df <- data.frame(
    year = c("1930",
             "1930",
             "1930"),
    month = c("12",
              "12",
              "12"),
    day = c("14",
            "15",
            "16"))

dates_df %>%
    unite(date, year, month, day, sep = "-")
```

### All together using the pipe

```{r}
mean_region <- catch_original %>% 
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>%  # Fix "I" in Chinook column
  mutate(Chinook = as.numeric(Chinook)) %>%  # Convert Chinook to numeric
  pivot_longer(-c(Region, Year), names_to = "species", values_to = "catch") %>%  # Reshape data
  mutate(catch = catch * 1000) %>%  # Multiply catch by 1000
  group_by(Region) %>%  # Group by Region
  summarise(mean_catch = mean(catch, na.rm = TRUE)) %>%  # Calculate mean catch
  arrange(desc(mean_catch))  # Sort by mean catch in descending order
```

